Canonical rings of rational divisors on projective space and Hirzebruch surfaces

### 2015-07-13 surfaces-write-up/surface-write-up.tex
lemma 5.3 edits

For $i \in T_=(D)$
define
\[
	\epsilon_i := (\prod_{k \ne i} f_k^{-\alpha \ell_i \gcd(\bida_i, \bidb_i)})
	(f_i^{\ell_i \frac{\gcd(\bida_i, \bidb_i)}{\bida_i + \bidb_i}\sum_{k \ne i}
	\alpha_k (\bida_k + \bidb_k)}) \in H^0(\ell_i \gcd(\bida_i, \bidb_i) D).
\]

For $i \in T_+(D)$ and $j \in T_-(D)$, define
\[
	\epsilon_{i, j} := (\prod_{k = 1}^m f_k^{-\alpha \ell_{i,j} (\bida_i \bidb_
j - \bida_j \bidb_i)}) {f_i}^{s_1} {f_j}^{s_2} \in H^0(\ell_{i,j}(\bida_i \bidb_j - 
\bida_j \bidb_i)D)
\]

where
\[
	s_1 = \bidb_j \sum_{k \ne i,j} \alpha_k \bida_k - \bida_j \sum_{k\ne i, j}
	\alpha_k \bidb_k
\]

\noindent
and
\[
	s_2 = \bidb_i \sum_{k \ne i,j} \alpha_k \bida_k - \bida_i \sum_{k \ne i, j}
	\alpha_k \bidb_k.
\]

@@

\begin{proof}
We first claim that $\psi$ has a relation of the form
\begin{equation}\label{eqn:relations-psi}
	(y_j - \beta_j)\prod_{i=1}^m {y_i}^{c_{i}}
\end{equation}

\noindent
for some $d\in \mathbb{N}$, $c_i \ge -\alpha_i d$, $\bida_j + \sum_{i = 1}
^m \bida_i c_i = 0$, $\bidb_j + \sum_{i=1}^m \bidb_i c_i = 0$, and $\beta_j \in
k[y_1, \ldots, y_4]$ satisfies $\psi(\beta_j)\in \mathscr{O}(\bida_j,
\bidb_j)$.

@@

Note that we can view the relations from Equation
\ref{eqn:relations-psi} as an element of $\mathbb{R}^{n+1}$ given
by $(d, c_1, \ldots, c_n)$.  Let $S'\subseteq \mathbb{R}^{+}$ be the
set of $(d, c_1, \ldots, c_n)$ satisfying $c_i \ge -\alpha_i d$,
$\bida_j + \sum_{i = 1}^m \bida_i c_i = 0$, and $\bidb_j + \sum_{i=1}^m \bidb_i c_i
= 0$.  Note that $S'$ is a translation of $S$ by \todo{finish}

>>>>

For $i \in T_=(D)$
define
\[
	\epsilon_i := (\prod_{k \ne i} f_k^{-\alpha_k \ell_i \gcd(\bida_i, \bidb_i)})
	(f_i^{\ell_i \frac{\gcd(\bida_i, \bidb_i)}{\bida_i + \bidb_i}\sum_{k \ne i}
	\alpha_k (\bida_k + \bidb_k)}) \in H^0(\ell_i \gcd(\bida_i, \bidb_i) D).
\]

For $i \in T_+(D)$ and $j \in T_-(D)$, define
\[
	\epsilon_{i, j} := (\prod_{k = 1}^m f_k^{-\alpha_k \ell_{i,j} (\bida_i \bidb_
j - \bida_j \bidb_i)}) {f_i}^{s_1} {f_j}^{s_2} \in H^0(\ell_{i,j}(\bida_i \bidb_j - 
\bida_j \bidb_i)D)
\]

where
\[
	s_1 = \bidb_j \sum_{k \ne i,j} \alpha_k \bida_k - \bida_j \sum_{k\ne i, j}
	\alpha_k \bidb_k = b_j (\alpha_i a_i + \alpha_j a_j) - a_j (\alpha_i b_i + \alpha_j b_j)
\]

\noindent
and
\[
	s_2 = -\bidb_i \sum_{k \ne i,j} \alpha_k \bida_k + \bida_i \sum_{k \ne i, j}
	\alpha_k \bidb_k = -b_i (\alpha_i a_i + \alpha_j a_j) + a_i (\alpha_i b_i + \alpha_j b_j).
\]

@@

\begin{proof}
We first claim that $\psi$ has a relation of the form
\begin{equation}\label{eqn:relations-psi}
	(y_j - \beta_j)\prod_{i=1}^n {y_i}^{c_{i}}
\end{equation}

\noindent
for some $d\in \mathbb{N}$, $c_i \ge -\alpha_i d$, $\bida_j + \sum_{i = 1}
^n \bida_i c_i = 0$, $\bidb_j + \sum_{i=1}^n \bidb_i c_i = 0$, and $\beta_j \in
k[y_1, \ldots, y_4]$ satisfies $\psi(\beta_j)\in \mathscr{O}(\bida_j,
\bidb_j)$.

@@

Note that we can view the relations from Equation
\ref{eqn:relations-psi} as an element of $S$ given
by $(d, c_1, \ldots, c_{i-1}, c_i + 1, c_{i+1}, \ldots , c_n)$.  Let $S'\subseteq (\mathbb{R}_{\ge 0})^{n+1}$ be the
set of $(d, c_1, \ldots, c_{i-1}, c_i + 1, c_{i+1}, \ldots , c_n)$ satisfying $c_i \ge -\alpha_i d$,
$\bida_j + \sum_{i = 1}^m \bida_i c_i = 0$, and $\bidb_j + \sum_{i=1}^n \bidb_i c_i
= 0$.  A careful computation deduces that $\gamma (d, c_1, \ldots, c_{i-1}, c_i + 1, c_{i+1}, \ldots, c_n)$ is of the form 
\[
	E + \sum_{k\in T_=} r_k e_k + \sum_{h\in T_+ \atop k\in T_-} r_{h,k} e_{h,k},
\] 
where $E$ is $\mathbb{R}_{\ge 0}$
combination of elements in 
\[
	\{e_i \mbox{ if } i\in S\} \cup \{e_{i,j}: j\in S_- \mbox{ if } i \in S_+\} \cup \{e_{j,i}: j\in S_+ \mbox{ if } i \in S_-\}
\]
given by lattice point $(d_E, w_1, \ldots, w_n)$ with $w_i = 1$.  These conditions imply that 
\[
	d_E \le \left\{ \begin{split}  \ell_i \gcd(\bida_i, \bidb_i) \mbox{ if } i \in S_= 
			\\ \max_{j\in S_-}(\ell_{i,j} (\bida_i \bidb_
j - \bida_j \bidb_i)) \mbox{ if } i\in S_+ \\
			\\ \max_{j\in S_+}(\ell_{j, i} (\bida_j \bidb_
i - \bida_i \bidb_j)) \mbox{ if } i \in S_- \end{split}  \right.
\]

Let $M$ be the set of monomials of $\beta_j$.  Then for $m\in M$, the corresponding lattice point is of the form
\[
	E_m + \sum_{k\in T_=} r_k e_k + \sum_{h\in T_+ \atop k\in T_-} r_{h,k} e_{h,k}
\]
where $E_m$ is a $\mathbb{R}_{\ge 0}$ combination of elements in
\[
	\{e_j : j\in T_= \cap \{1, 2, 3, 4\}\} \cup \{e_{j, k}: j\in T_= \cap \{1, 2, 3, 4\}, k\in T_-\} \cup \{e_{k,j}: k\in T_+, j\in T_- \cap \{1, 2, 3, 4\}\}.
\]
